---
title: "Statisical Inference Course Project - Part 2"
author: "Alex McBride"
date: "Saturday, February 07, 2015"
output: 
      html_document:
      keep_md: yes
      toc: yes
---

## Summary of Project

Analyze the ToothGrowth data in the R datasets package. Load the ToothGrowth data and perform some basic exploratory data analyses. Provide a basic summary of the data. Use confidence intervals and/or hypothesis tests, as taught in class to compare tooth growth by "supp" and "dose". Finally, state your conclusions and the assumptions needed for your conclusions. 

## Load the data and libraries
```{r, echo=TRUE}
library(ggplot2)
library(grid)
data("ToothGrowth")
data <- as.data.frame(ToothGrowth)
summary(data)
```

```{r}
# Get the mean length per dose group

dataag <- aggregate(len~dose+supp, data=data, mean)

```


```{r}
# Create Plots
p1 <- ggplot(data, aes(x=factor(supp), y=len)) + 
           geom_bar(stat="identity", aes(fill=factor(dose)), position = "dodge") +
              scale_fill_brewer(name="Dose") +
           theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
           guides(fill = guide_legend(reverse = TRUE)) +
           labs(x="Supplement", y="Length",
                title = "Maximum Tooth Growth")                  

p2 <- ggplot(dataag, aes(x=factor(supp), y=len)) + 
     geom_bar(stat="identity", aes(fill=factor(dose)), position = "dodge") +
        scale_fill_brewer(name="Dose") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
     guides(fill = guide_legend(reverse = TRUE)) +
     labs(x="Supplement", y="Average Length Grown",
          title = "Mean Tooth Growth")

# Define layout for plots and print
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
```

